---
title: "model_result_analysis"
output: html_document
---

Just Museumstation Tracing (no photodraw data)

```{r}
library(tidyverse)
library(lme4)
library(groupdata2)
library(dplyr)
library(knitr) 
library(broom) 
library(hydroGOF) 
```

Only Tracings: shape + square
```{r}
raw = read.csv('museumstation_tracing_eval.csv')
raw[raw$human_norm_rating==-10.0, 'human_norm_rating'] = NA
raw[raw$human_rating==-10.0, 'human_rating'] = NA
raw$has_ref = as.logical(raw$has_ref)
raw$norm_r = scale(raw$rotate)
raw$norm_t = scale(raw$translate)
raw$norm_s = scale(raw$scale)
raw$spatial = raw$norm_r + raw$norm_t + raw$norm_s
raw$norm_shape = scale(raw$post_tran)
raw$norm_spatial = scale(raw$spatial)
summary(raw)
```
```{r}
d = raw[raw$has_ref==TRUE,]
copy = raw[raw$has_ref==FALSE,]
```

#### plot shape error and spatial error
##### Tracing
```{r}
require(gridExtra)
square = d[d$category == 'square',]
p_square = ggplot(square, aes(norm_shape, norm_spatial, color=age)) +
  geom_point(size=1) +
  theme_bw() +
  ggtitle('Square Tracing') + 
  xlab("norm_shape") +
  ylab("norm_spatial")

shape = d[d$category == 'shape',]
p_shape = ggplot(shape, aes(norm_shape, norm_spatial, color=age))+
  geom_point(size=1) +
  theme_bw() + 
  ggtitle('Shape Tracing') + 
  xlab("norm_shape") + 
  ylab("norm_spatial")

grid.arrange(p_square, p_shape, ncol=2)
```
##### Copying
```{r}
square = copy[copy$category == 'square',]
p_square = ggplot(square, aes(norm_shape, norm_spatial, color=age)) +
  geom_point(size=1) +
  theme_bw() +
  ggtitle('Square Copying') + 
  xlab("norm_shape") +
  ylab("norm_spatial")

circle = copy[copy$category == 'circle',]
p_circle = ggplot(circle, aes(norm_shape, norm_spatial, color=age))+
  geom_point(size=1) +
  theme_bw() + 
  ggtitle('Circle Copying') +
  xlab("norm_shape") + 
  ylab("norm_spatial")

grid.arrange(p_square, p_circle, ncol=2)
```

#### correlation
```{r}
cor.test(raw$norm_shape, raw$norm_spatial)
```

##### Tracing: shape + square
```{r}
cor.test(d$norm_shape, d$norm_spatial)
```

##### Copying: square + circle
```{r}
cor.test(copy$norm_shape, copy$norm_spatial)
```

#### Model Fitting (only for tracing)
For tracing, is norm_shape like some kind of threshold for the norm_spatial?

```{r}
m1 = lm('norm_spatial ~ norm_shape + category + age', d)
summary(m1)
```
```{r}
m2 = lm('norm_shape ~ norm_spatial + category + age', d)
summary(m2)
```

##### For the rated tracings
```{r}
rdata = d[!is.na(d$human_rating),]
cor.test(rdata$norm_shape, rdata$norm_spatial)
```

```{r}
m1 = lm('norm_spatial ~ norm_shape + category + age', rdata)
summary(m1)
```

```{r}
m2 = lm('norm_shape ~ norm_spatial + category + age', rdata)
summary(m2)
```

##### plot the change of shape_error and spatial_error as a function of age
```{r}
require(gridExtra)
for (x in c('square', 'shape')){
  a_shape = ggplot(d[d$category==x,], aes(age, norm_shape, color=norm_spatial))+
    geom_point(size=1) + 
    theme_bw() + 
    xlab("age") + 
    ylab("norm_shape")
  
  a_spatial = ggplot(d[d$category==x,], aes(age, norm_spatial, color=norm_shape)) + 
    geom_point(size=1) + 
    theme_bw() + 
    xlab("age") + 
    ylab("norm_spatial")
  grid.arrange(a_shape, a_spatial, ncol=2)
}

```

pre-transformation shape error vs. spatial error
```{r}
p = ggplot(d, aes(pre_tran, norm_spatial, color=age)) + 
  geom_point(size=1) + 
  theme_bw() + 
  xlab('pre-transformation shape error') + 
  ylab('spatial error')
print (p)
```
