
---
title: "Correlation between Tracing Analysis Mturk Human rating"
output: html_document
---

```{r}
library(tidyverse)
library(lme4)
library(groupdata2)
library(dplyr)
library(knitr) # kable()
library(broom) #tidy()
library(hydroGOF) # rmse()
```

### Summary

#### Basic information about the survey
In total: 80 stimuli per shape per age - 2160 stimuli

Current data: 1982 stimuli get rated by 26 mturkers (1 mturker only pressed 4 and his/her answers are removed)

#### Normalized Cross-Correlation Loss

For a reference and a drawing:

1. Multiply the value (0-1) at each pair of corresponding pixels and sum them up

2. Normalize the sum to mean=0, std=1

#### Import rater_data
```{r}
rater_data <- read.csv('museumstation_tracing_ncc.csv')%>%
  mutate(tracing_item = factor(tracing_item, levels = c('square', 'shape', 'circle')))
```

#### Current data in each age + shape group
```{r}
for (t in c('square', 'shape', 'circle')){
  for (i in seq(2, 10, by=1) ){
    sub = subset(rater_data , tracing_item == t & age == i)
    print (c(t, i, nrow(sub)) )
  }
}
```


```{r}
rater_data$overall = -1 * scale(rater_data$overall)

rater_data %>%
  gather(registration, val, human_norm_rating, overall) %>%
  ggplot(aes(x = age, y = val, color = tracing_item)) +
    geom_point() +
    geom_smooth(method = 'lm') +
    facet_wrap(~ registration) +
    theme_bw() +
    theme(aspect.ratio = 1) 
```

```{r}
rater_data$norm_shape = -1 * scale(rater_data$norm_shape)
rater_data %>%
  gather(registration, val, human_norm_rating, norm_shape) %>%
  ggplot(aes(x = age, y = val, color = tracing_item)) +
    geom_point() +
    geom_smooth(method = 'lm') +
    facet_wrap(~ registration) +
    theme_bw() +
    theme(aspect.ratio = 1) 
```

```{r}
rater_data$norm_spatial = -1 * scale(rater_data$norm_spatial)
rater_data %>%
  gather(registration, val, human_norm_rating, norm_spatial) %>%
  ggplot(aes(x = age, y = val, color = tracing_item)) +
    geom_point() +
    geom_smooth(method = 'lm') +
    facet_wrap(~ registration) +
    theme_bw() +
    theme(aspect.ratio = 1) 
```

#### Corrleation of human rating with NCC Loss: 
Shape Error vs. Human Rating: r = - 0.70, p < 0.01

Spatial Error vs. Human Rating: r = - 0.51, p < 0.01

Shape+Spatial Error vs. Human Rating: r = - 0.69, p < 0.01

```{r}
rater_data <- read.csv('kiddraw_tracing_summary.csv')%>%
  mutate(tracing_item = factor(tracing_item, levels = c('square', 'shape', 'circle')))
cor.test(rater_data$norm_shape, rater_data$human_norm_rating)
```
```{r}
cor.test(rater_data$norm_spatial, rater_data$human_norm_rating)
```
```{r}
cor.test(rater_data$overall, rater_data$human_norm_rating)
```

#### Cross-Validation

```{r}
mdata <- read.csv('museumstation_tracing_ncc.csv')
#rater_data$copy = 0
#rater_data[rater_data$tracing_item == 'circle',]$copy = 1
#rater_data$copy = factor(rater_data$copy, levels = c(0,1))
mdata$tracing_item = factor(mdata$tracing_item, levels = c('square', 'shape', 'circle'))
rater_data = mdata[mdata$human_rating!=-10.0,]
```

```{r}
# Split data in 20/80 test/training
set.seed(1)
parts <- partition(rater_data, p = 0.2, cat_col = c("age", "tracing_item"))

test_set <- parts[[1]]
train_set <- parts[[2]]
```

```{r}
train_set <- fold(train_set, k=5, cat_col = c("age", "tracing_item"))

train_set <- train_set[order(train_set$.folds),]

# train_set %>% 
#   count(age, tracing_item, .folds) %>% 
#   kable(align='c')

```

```{r}
crossvalidate <- function(data, k, model, dependent, random = FALSE){
  performances <- c()
  for (fold in 1:k){
    testing_set <- data[data$.folds == fold,]
    training_set <- data[data$.folds != fold,]
    #print (nrow(testing_set))
    #print (nrow(training_set))

    if (isTRUE(random)){

      # Train linear mixed effects model on training set
      model <-  lmer(model, training_set, REML=FALSE)

    } else {

      # Train linear model on training set
      model <-  lm(model, training_set)

    }

    ## Test model

    # Predict the dependent variable in the training_set with the trained model
    predicted <- predict(model, testing_set, allow.new.levels=TRUE)
    MSE <- mean((predicted - testing_set[[dependent]])^2)
    
    # Predict the dependent variable in the test_set with the trained model
    predicted_train <- predict(model, training_set, allow.new.levels=TRUE)
    MSE_train <- mean((predicted_train - training_set[[dependent]])^2)
    
    performances[fold] <- MSE
    print (paste0(fold, "  training MSE: ", MSE_train, "  hold out MSE: ", MSE))
    
    test_r = (predicted - testing_set[[dependent]])^2
    train_r = (predicted_train - training_set[[dependent]])^2
    
    ## Plot the fold
    # training_set$model = predicted_train
    # testing_set$model = predicted
    # 
    # p = ggplot(training_set, aes(model, human_norm_rating, color=tracing_item)) +
    #   geom_point(size=3) +
    #   geom_abline(intercept = 0, slope = 1) + 
    #   theme_bw() + 
    #   xlab("model") +
    #   ylab("human") 
    # print (p)
    # 
    # p = ggplot(testing_set, aes(model, human_norm_rating, color=tracing_item)) +
    #   geom_point(size=3) +
    #   geom_abline(intercept = 0, slope = 1) + 
    #   theme_bw() + 
    #   xlab("model") +
    #   ylab("human") 
    # print (p)  
  
  }

  return (performances)

}
```

#### Shape Error + Spatial Error
Model: Shape_Error + Spatial_Error + Age + Tracing_Item + (1 | Session_ID)
```{r}
m1 = "human_norm_rating ~ age + norm_shape + norm_spatial + tracing_item + (1 | session_id)"
result = crossvalidate(train_set, k=5, model=m1, dependent='human_norm_rating', random=TRUE)
print (paste0("Average MSE across all folds: ", mean(result)))
```

#### Spatial Error Only
Model: Spatial_Error + Age + Tracing_Item + (1 | Session_ID)
```{r}

m4 = "human_norm_rating ~ age + norm_spatial + tracing_item + (1 | session_id)"
result = crossvalidate(train_set, k=5, model=m4, dependent='human_norm_rating', random=TRUE)
print (paste0("Average MSE across all folds: ", mean(result)))

```

#### Shape Error Only

Model: Shape_Error + Age + Tracing_Item + (1 | Session_ID)
```{r}
m5 = "human_norm_rating ~ age + norm_shape + tracing_item + (1 | session_id)"
result = crossvalidate(train_set, k=5, model=m5, dependent='human_norm_rating', random=TRUE)
print (paste0("Average MSE across all folds: ", mean(result)))
```

##### Apply the shape + spatial error model on the whole training dataset and examine it on the new test set
Model: Shape_Error + Spatial_Error + Age + Tracing_Item + (1 | Session_ID)
```{r}
# Creating the model for the full training set
model1 <- lmer(m1, train_set, REML = FALSE)
# Predict the dependent variable in the test_set with the trained model
predicted <- predict(model1, test_set, allow.new.levels=TRUE)
# Get the Root Mean Square Error between the predicted and the observed
MSE <- mean((predicted - test_set[['human_norm_rating']])^2)

summary(model1)
MSE
```

#### Shape Error + Spatial Error + Interaction between Shape and Spatial

- When shape error is large, participants might give the tracing a low score despite the spatial error.
- When shape error is small, participants might starts to weight spatial error more heavily.


Model: Shape_Error + Spatial_Error +  Shape_Error * Spatial_Error + Age + Tracing_Item + (1 | Session_ID)
```{r}
#m6 = "human_norm_rating ~ age + norm_shape + norm_spatial + norm_shape * norm_spatial + tracing_item + (1 | session_id)"
m6 = "human_norm_rating ~  copy + tracing_item + norm_shape * norm_spatial"
result = crossvalidate(train_set, k=5, model=m6, dependent='human_norm_rating', random=FALSE)
print (paste0("Average MSE across all folds: ", mean(result)))
```

##### Apply the interaction model on the whole training dataset and examine it on the new test set
Model: Shape_Error + Spatial_Error +  Shape_Error * Spatial_Error + Age + Tracing_Item + (1 | Session_ID)
```{r}
# Creating the model for the full training set
inter <- lm(m6, train_set, REML = FALSE)
# Predict the dependent variable in the test_set with the trained model
predicted <- predict(inter, test_set, allow.new.levels=TRUE)
# Get the Root Mean Square Error between the predicted and the observed
MSE <- mean((predicted - test_set[['human_norm_rating']])^2)

summary(inter)
MSE
```

```{r}
anova(inter)
```

```{r}
#hist(test_r, col="gray", labels = TRUE, xlim=c(0,4))
    #hist(train_r, col="gray", labels = TRUE, xlim=c(0,4))
    #plot(predicted_train, training_set[[dependent]])
    #plot(predicted, testing_set[[dependent]])
    
```

```{r}
raw = read.csv('museumstation_tracing_ncc2.csv')
#raw = read.csv('tracing_ordinal_data.csv')
raw$category = factor(raw$category, levels = c('square', 'shape', 'circle'))
raw$has_ref = as.logical(raw$has_ref)
raw = raw[raw$has_ref==TRUE,]
raw[raw$human_norm_rating==-10.0, 'human_norm_rating'] = NA
raw[raw$human_rating==-10.0, 'human_rating'] = NA
write.csv(raw, 'tracing_raw.csv', row.names = FALSE)
```

```{r}
class = read.csv('classification.csv')
```

